Decentralized autonomous organization to automate governance\\ {\smaller \textbf{Final Draft - UNDER REVIEW}}

### 2016-04-14 paper/Paper.tex

Historically, corporations have only been able to act through people (or through corporate entities that were themselves ultimately controlled by people). This presents two simple and fundamental problems. Whatever a private contract or public law require:  (1) people do not always follow the rules and (2) people do not always agree what the rules actually require. Collaboration without a corporate form does not solve these problems, necessarily, and it may introduce others. In the absence of a coporate form, an explicit written agreement is substituted for unclear informal ``understandings'' and legal protections provided by a corporate form will not be available.

@@

Crowdfunding (\cite{2015CFReport}) amplifies the problem. On the one hand, it has made it easier for small contributors to invest in large projects. It has also made it possible for entrepeneurs to receive financial support that might not have been easily available in the past. On the other hand, small investors remain vulnerable to financial mismanagement or outright fraud. Because they have small stakes in a venture, they may lack power to identify problems, participate in governance decisions, or to easily recover their investment (\cite{9MostDisgracefulCrowdFundings, CrowdfundingFailTC}). At the same time, corporate leadership and managment may be accused of malfeasance or mismanagement when they believe that they have acted in good faith, based on their understanding of their obligations and interpretation of applicable rules.

We present a potential solution using Ethereum, (\cite{buterin2013ethereum,Wood2014ethereum}) a blockchain technology which integrates a Turing complete programming language with smart contract processing functionality. We illustrate a method that for the first time allows the creation of organizations in which (1) participants maintain direct real-time control of contributed funds and (2) where governance rules are formalized and automated using software. Specifically, we have written generic smart contract code (\cite{szabo1997formalizing, miller1997future}) that can be used to form a Decentralized Autonomous Organization (DAO) on the Ethereum blockchain. In this paper we explain how the DAO code works. We do so by focusing on some basic formation and governance features, including formation, funding and voting rights.

We first describe the DAO's formation and basic functionality. We then discuss minority owner rights and propose a solution to the ``Majority Robbing the Minority Attack'' problem, the ``DAO split''. We then explore the smart contract code in detail, and conclude with an explanation and detailed specification of the ``DAO split''.

@@

Once deployed, the DAO code requires ``ether'' to engage in transactions on Ethereum. Ether is the digital fuel that powers the Ethereum network. Without ether, a DAO can not do very much and a DAO's first order of business is to receive ether. After the DAO code is deployed, ether may be sent to the DAO's smart contract address during an initial funding period which is defined in the DAO code.

In exchange for ether, the DAO code creates tokens that are then assigned to the account of the person who sent the ether. A token grants its holder voting and ownership rights. The number of tokens created is proportional to the amount of ether transferred. Token price varies over time (see section \ref{TokenPrice}). Token ownership is freely transferable on the Ethereum blockchain, when the funding period is over.

A minimum DAO funding goal and funding time-period are set as parameters in the DAO code during to deployment. If the minimum funding amount is not reached at the close of the funding period, all ether is returned. After the funding period has ended, we denote the total ether raised by $\Xi_{\text{raised}}$ and the total amount of tokens created by $T_{\text{total}}$.

The DAO stores ether and transmits it based on the DAO code. It does not do much else. It cannot build a product, write code or develop hardware. It requires what we have called a ``Service Provider'' to accomplish these and other goals. The DAO engages a Service Provider by accepting a Service Provider proposal.

DAO token holders may also submit proposals to use the DAO's ether, denoted by $\Xi_{\text{transfer}}$, for a variety of purposes. If a proposal is approved, the DAO transmits ether to a smart contract representing the proposed project. Such smart contracts can be parameterized and enable the DAO to interact with and influence the project being funded.
An example of such an agreement between the DAO and a project to be funded can be found in the appendix (\ref{app:SampleOffer}).

Members of the DAO cast votes weighted by the amount of tokens they control. Tokens are divisible, indistinguishable, fungible and can easily be transferred between accounts. Within the contracts, the individual actions of members, cannot be directly determined. There is a set timeframe $t_p$ to debate and vote on any given proposal. In our example, this timeframe is set by the creator of the proposal, and is required to be at least two weeks for a regular proposal.

After $t_p$ has passed, any token holder can call a function in the DAO contract that will verify that the majority voted in favor of the proposal and that quorum was reached; the function will execute the proposal if this is the case. If this is not the case, the proposal will be closed. The minimum quorum represents the minimum number of tokens required for a vote to be valid, is denoted by $q_{min}$, and calculated as follows:
\begin{equation}
 q_{min} = \frac{T_{\text{total}}}{d} + \frac{\Xi_{\text{transfer}} \cdot T_{\text{total}}}{3 \cdot (\Xi_{DAO} + R_{\text{DAO}})}
 \label{minQuorum}
\end{equation}
Where $d$ is the \verb|minQuorumDivisor|. This parameter's default value is $5$, but it will double in the case the quorum has not been met for over a year.
$\Xi_{DAO}$ is the amount of ether owned by the DAO and $R_{\text{DAO}}$ is the amount of reward tokens owned by this DAO, as explained in section \ref{RewardToken} (also see \verb|rewardToken| in \ref{app:DAO}). The sum $\Xi_{DAO} + R_{\text{DAO}}$ is equal to the amount of ether raised plus the rewards received or said another way, the total amount of ether the DAO has ever received.

This means, initially, a quorum of  $20\%$ of all tokens is required for any proposal to pass. In the event $\Xi_{\text{transfer}}$ equals the amount of ether raised during the funding period plus the rewards received, then a quorum of $53.33\%$ is required.

In order to prevent ``proposal spam'', a minimal deposit is required to be paid when creating a proposal, which gets refunded if quorum is achieved. If quorum is not achieved, the DAO keeps the proposal deposit. The value of the proposal deposit can be changed from the default value by the DAO through another proposal.

\section{Notation}
Throughout this paper, $\Xi$ always represents an amount of ether in its base unit wei. This is defined as $1 \: \text{Wei} = 10^{-18} \: \text{Ether}$ (\cite{Wood2014ethereum}).
Similarly, DAO tokens are denoted with $T$ and always represent the amount of DAO tokens in its base unit. We define this as $10^{-16}$ DAO token.
\section{Majority robs minority attack}
Minority owner rights can be a problem in any corporate form. Minority rights may be protected or addressed by provisions in corporate governance documents or by statute or judge-made law. But many of these solutions fail because minority owners may lack voting rights or the ability to ``vote with their feet" and easily retrieve their capital. We present a solution to this problem in the DAO code.

@@

To prevent this, the minority must always have the ability to retrieve their portion of the funds. Our solution is to allow the DAO to split into two. If an individual, or a group of token holders, disagree with a proposal and want to retrieve their portion of the funds before the proposal gets executed, they can submit a special type of proposal to form a new DAO. The minority can then vote to move their funds to this new DAO, leaving the majority alone only able to spend their own money.
This idea originates from a blog post by Vitalik Buterin (\cite{Vitalik2015subjectivity}).

A problem this simple fix doesn't address is voter apathy: some token holders might not be actively involved in the DAO and might not follow proposals closely.
An attacker could use this to their advantage. Even though the minority has the chance to retrieve their funds and split the DAO, some could be unaware of the situation and fail to act.
For the DAO to be considered safe, it is required that inactive token holders must also be protected from losing their ether. Our proposed solution is implemented by limiting each individual DAO to a single curator.
This curator controls the list of addresses that can receive ether from the DAO, across all proposals.
This gives the curator of the DAO considerable power. To prevent the abuse of this power, it is possible for the DAO to vote for a new curator, which may result in a split of the DAO as described above.

Any token holder can make a proposal to vote for a new curator. In effect, even a single token holder is able to both retrieve their remaining portion of ether and maintain their right to any future rewards associated to their previous contribution, as these will be sent to the new DAO automatically. Rewards are defined as any ether received by the DAO generated from products the DAO funded so far and are explained in further detail in section \ref{RewardToken}.

The process of choosing a new curator is as follows:
Any token holder can submit a proposal for a new curator. In this case, no proposal deposit is required, because an attacker could vote for an extremely high deposit, preventing any splits. The debating period for this proposal is $7$ days.
This is $7$ days less than the minimum required for regular proposals, allowing anyone to retrieve their funds before a potentially malicious proposal goes through. There is no quorum requirement, so that every token holder has the ability to split into their own DAO. 
The debating period is used to discuss (on or off-chain) the new curator and conduct a first vote that's non-binding. After this first vote, token holders can cast a second one to confirm its results. In the case a majority opts to  keep the original curator, the minority can either change their vote to keep the original curator in order to avoid a split, or inversely they can confirm their vote for a new curator and move their portion of the funds to a new DAO.

\section{Token Price} \label{TokenPrice}
DAO Token price increases over time. This reflects an assumption that early purchasers have greater risk, as they may have less information about the potential success of the DAO and do not know whether follow-on contribution will lead to full funding of the DAO. 
For the DAO described here, we have chosen the following price schedule:
\begin{equation}
 P(t) = \begin{cases}
1 & \text{if} \quad t < t_c - 2 \cdot \text{w} \\
1 + 0.05 \cdot m(t) & \text{if} \quad t_c - 2 \cdot \text{w} \leqslant t < t_c - 4 \cdot \text{d} \\
1.5 & \text{otherwise}
\end{cases}
\end{equation}
with the multiplier $m$ defined as:
\begin{equation}
 m(t) = (t - (t_c - 2 \cdot \text{w})) / \text{d}
\end{equation}
Here $t$ is the unix time in seconds, $t_c$ is the closing time of the funding period (see \ref{app:TokenSale} \verb|closingTime|), $w$ is a week in seconds and $d$ a day in seconds.
Hence the number of tokens (in it’s base unit) each buyer receives is calculated as: $P(t) \cdot \Xi_c$. Here $\Xi_c$ stands for the amount of ether payed, denoted in wei.
This results in a constant price in the beginning, until $2$ weeks before the end of the token sale. At this time the price increased daily by $0.05 \: \Xi_c$ per base unit of DAO token. Until $4$ days before the closing time when there will be a constant price of $1.5 \: \Xi_c$ per base unit of DAO token.

Price increases during the initial funding period could lead to a situation where a single contributor, having bought tokens during the initial price, could split the DAO immediately after the sale ends, thereby receiving more ether than they put in due to other contributors paying a higher price (\cite{GriffDiscussion}).
In order to avoid that possibility, all ether that is payed above the initial price of the token, will be sent to an extra account. Denoted as \verb|extraBalance| in \ref{app:TokenSale}. This money can be sent back to the DAO through a proposal after the DAO has spent at least this amount of ether.
This rule is implemented in the internal function \verb|isRecipientAllowed| in section \ref{DAO}.

@@

The address \verb|curator| is set at the creation of the DAO and defines the curator.

@@

The integer \verb|proposalDeposit| specifies the minimum deposit to be paid in wei for any proposal that does not include a change of the curator.

@@

A single proposal has the parameters:
\begin{description}
 \item[recipient] The address where the \verb|amount| of wei will go to if the proposal is accepted.
 \item[amount] The amount of wei to transfer to \verb|recipient| if the proposal is accepted.
 \item[description] A plain text description of the proposal.
 \item[votingDeadline] A unix timestamp, denoting the end of the voting period.
 \item[open] A boolean which is false if the votes have already been counted, true otherwise.
 \item[proposalPassed] A boolean which is true if a quorum has been achieved with the majority approving the proposal.
 \item[proposalHash] A hash to check validity of a proposal. Equal to \verb|sha3(_recipient, _amount, _transactionData)|.
 \item[proposalDeposit] The deposit (in wei) the creator of a proposal has send to submit a proposal. It is taken from the \verb|msg.value| of a \verb|newProposal| call; its purpose is to prevent spam. Its minimal value is set when the contract is deployed as constructor parameter. But the creator of the proposal can send any amount above this for the deposit. The proposals will be sorted by the proposal deposit in the GUI, so in the case that a proposal is considered important, the creator of the proposal can deposit extra ether to “advertise” their proposal. The creator of the proposal will be refunded the entire deposit if quorum is reached, if it is not reached the deposit remains with the DAO. 
 \item[newCurator] A boolean which is true if this proposal assigns a new curator.
 \item[splitData] The data used to split the DAO. This data is gathered from proposals when they require a new curator.
 \item[yea] Number of tokens in favor of the proposal.
 \item[nay] Number of tokens opposed to the proposal.
 \item[votedYes] Simple mapping to check if a token holder has voted for it.
 \item[votedNo] Simple mapping to check if a token holder has voted against it.
 \item[creator] The address of the token holder that created the proposal.
\end{description}

@@

\subsubsection*{newProposal}
This function is used to create a new proposal. The arguments of the function are:
\begin{description}
 \item[recipient] The address of the recipient of the ether in the proposal (has to be the DAO address itself, the current curator or an address on the whitelist \verb|allowedRecipients|).
 \item[amount] The amount of wei to be sent in the proposed transaction.
 \item[description] A string describing the proposal.
 \item[transactionData] The data of the proposed transaction.
 \item[debatingPeriod] The amount of time to debate the proposal, at least 2 weeks for a normal proposal and at least 1 week for a new curator proposal.
 \item[newCurator] A boolean defining whether this proposal is for a new curator or not.
\end{description}
After checking the sanity of the proposal (see code), this function creates a proposal which is open for voting for a certain amount of time. The function will return a proposal ID which is used to vote.

@@

\subsubsection*{executeProposal}
This function can be called by anyone. It counts the votes, in order to check whether the quorum is met, and executes the proposal if it passed, unless it is a proposal for a new curator, than it does nothing.
The arguments of the function are:
\begin{description}
 \item[proposalID] The proposal ID.
 \item[transactionData] The data of the proposed transaction
\end{description}
The function checks whether the voting deadline has passed and that the \verb|transactionData| matches the proposal ID. Then it checks whether the quorum has been met (see Eq. \ref{minQuorum}) and if the proposal had a majority of support. If this is the case, it executes the proposal and refunds the 
proposal deposit. If the quorum has been achieved, but the proposal was declined by the majority of the voters, the proposal deposit is refunded and the proposal closes.

\subsubsection*{emergenyCloseProposal}
This is used in the case `executeProposal` fails because it throws (e.g. when returuning the deposit, because the fallback function of this address throws, or when executing the proposed transaction). It can be called $5$ days after the voting deadline of the proposal passed. It will close the proposal.
\subsubsection*{splitDAO}
After a new curator has been proposed, and the debating period in which the token holders could vote for or against the proposal has passed, this function is called by each of the DAO token holders that want to leave the current DAO and move to a new DAO with the proposed new curator. This function creates a new DAO and moves a portion of the ether, as well as a portion of the reward tokens to the new DAO.
The arguments are:
\begin{description}
 \item[proposalID] The proposal ID.
 \item[newCurator] The address of the new curator of the new DAO.
\end{description}
After a sanity check (see code), this function will create the new DAO if it hasn’t already been created using the contract \verb|daoCreator|, updates the split data stored in the proposal and stores the address of the new DAO in the split data.
This function moves the portion of ether that belongs to the caller of this function in the original DAO to the new DAO. This ether amount is denoted by $\Xi_{\text{sender}}$, stated in wei and is calculated as follows:
\begin{equation}
 \Xi_{\text{sender}} = \Xi_{\text{DAO}} \cdot T_{\text{sender}} / T_{\text{total}}
\end{equation}
Here $T_{\text{sender}}$ is the amount of tokens of the caller of the function and $\Xi_{\text{DAO}}$ is the balance of the DAO at the time of the split.
This will be used to effectively buy tokens of the newly created DAO and fund the new DAO just as the original DAO was funded.
In addition to the ether which is moved to the new DAO, the reward tokens $R_{\text{sender}}$ are also transferred. 
They are calculated as follows:
\begin{equation}
 R_{\text{sender}} = R_{\text{DAO}} \cdot T_{\text{sender}} / T_{\text{total}}
\end{equation}
Where $R_{\text{DAO}}$ is the amount of reward tokens owned by the original DAO at the time of the split. These tokens allow the new DAO to retrieve their portion of the reward using the \verb|retrieveDAOReward| function of the original DAO.
At the end of this process all original DAO tokens of the sender account are destroyed. It is important to notice that in all integer division descirbed above, there may be remainders which stay with the DAO.

\subsubsection*{newContract}
This function can only be called by the DAO itself (through a proposal and a vote) and is used to move all remaining ether, as well as all rewardTokens to a new address. This is used to update the contract. The new address needs to be approved by the curator.

@@

\subsubsection*{changeAllowedRecipients}
This function can add/remove an address to/from the whitelist, \verb|allowedRecipients|. It can only be executed by the curator.

@@

Let’s assume that during the split, a fraction of DAO tokens, $X$, changes the curator and leaves the DAO. The new DAO created receives $X \cdot \Xi_{\text{DAO pre}}$, a portion of the remaining ether from the original DAO.
\begin{equation}
 \Xi_{\text{DAO post}} = (1 - X) \cdot \Xi_{\text{DAO pre}}
\end{equation}
Here $\Xi_{\text{DAO pre}}$ is the ether balance of the original DAO before the split and $\Xi_{\text{DAO post}}$ is the ether balance of the original DAO after the split.

@@

The original DAO tokens of the accounts that confirmed the new curator are destroyed.
Hence:
\begin{equation}
 T_{\text{total post}}= (1 - X) \cdot T_{\text{total pre}}
\end{equation}

@@

    // Address of the curator
    address public curator;
    // The whitelist: List of addresses the DAO is allowed to send money to
    mapping (address => bool) public allowedRecipients;

>>>>

Historically, corporations have only been able to act through people (or through corporate entities that were themselves ultimately controlled by people). This presents two simple and fundamental problems. Whatever a private contract or public law require:  (1) people do not always follow the rules and (2) people do not always agree what the rules actually require. Collaboration without a corporate form does not solve these problems, necessarily, and it may introduce others. In the absence of a corporate form, an explicit written agreement is substituted for unclear informal ``understandings'' and the legal protections provided by a corporate form will not be available.

@@

Crowdfunding (\cite{2015CFReport}) amplifies the problem. On the one hand, it has made it easier for small contributors to invest in large projects, and it has also made it possible for entrepreneurs to receive financial support that might not have been easily available in the past. On the other hand, small investors remain vulnerable to financial mismanagement or outright fraud, and because they have a small stake in a venture, they may lack power to identify problems, participate in governance decisions, or to easily recover their investment (\cite{9MostDisgracefulCrowdFundings, CrowdfundingFailTC}). At the same time, corporate leadership and management may be accused of malfeasance or mismanagement when they believe that they have acted in good faith, based on their understanding of their obligations and interpretation of applicable rules.

We present a potential solution using Ethereum, (\cite{buterin2013ethereum,Wood2014ethereum}) a blockchain technology which integrates a Turing complete programming language with smart contract processing functionality. We illustrate a method that for the first time allows the creation of organizations in which (1) participants maintain direct real-time control of contributed funds and (2) where governance rules are formalized, automated and enforced using software. Specifically, we have written generic smart contract code (\cite{szabo1997formalizing, miller1997future}) that can be used to form a Decentralized Autonomous Organization (DAO) on the Ethereum blockchain. In this paper we explain how a DAO's code works. We do so by focusing on some basic formation and governance features, including formation, creation and voting rights.

We first describe a DAO's formation and basic functionality. We then discuss minority owner rights and propose a solution to the ``Majority Robbing the Minority Attack'' problem, the ``DAO split''. We then explore the smart contract code in detail, and conclude with an explanation and detailed specification of the ``DAO split''.

@@

Once deployed, a DAO's code requires ``ether'' to engage in transactions on Ethereum. Ether is the digital fuel that powers the Ethereum network. Without ether, a DAO can not do anything so a DAO's first order of business is to receive ether. After a DAO's code is deployed, ether may be sent to the DAO's smart contract address during an initial Creation Phase which is defined in the DAO's code.

In exchange for ether, a DAO's code creates tokens that are assigned to the account of the person who sent the ether. The token grants its holder voting and ownership rights. The number of tokens created is proportional to the amount of ether transferred. Token price varies over time (see section \ref{TokenPrice}). Token ownership is freely transferable on the Ethereum blockchain, when the Creation Phase has ended.

A minimum DAO Creation goal and Creation Phase time-period are set as parameters in a DAO's code at the time of deployment. If the minimum DAO Creation goal is not reached at the close of the Creation Phase, all ether is returned. After the Creation Phase has ended, the total ether raised is denoted by $\Xi_{\text{raised}}$ and the total amount of tokens created is denoted by $T_{\text{total}}$.

A DAO stores ether and other Ethereum based tokens and transmits them based on the DAO's code. It does not do much else. It cannot build a product, write code or develop hardware. It requires what we have called a ``Contractor'' to accomplish these and other goals. A DAO selects a Contractor by accepting a Contractor's proposal.

Any DAO Token Holder may become a Contractor by submitting proposals to use a DAO's ether, denoted by $\Xi_{\text{transfer}}$. If a proposal is approved, the DAO transmits ether to a smart contract representing the proposed project. Such smart contracts can be parameterized and enable a DAO to interact with and influence the project it chose to support.
An example of such an agreement between a DAO and a project to be funded can be found in the appendix (\ref{app:SampleOffer}).

Members of a DAO cast votes weighted by the amount of tokens they control. Tokens are divisible, indistinguishable and can easily be transferred between accounts. Within the contracts, the individual actions of members, cannot be directly determined. There is a set time frame $t_p$ to debate and vote on any given proposal. In our example, this time frame is set by the creator of the proposal, and is required to be at least two weeks for a regular proposal.

After $t_p$ has passed, any token holder can call a function in the DAO contract that will verify that the majority voted in favor of the proposal and that quorum was reached; the function will execute the proposal if this is the case. If this is not the case, the proposal will be closed.

The minimum quorum represents the minimum number of tokens required for a vote to be valid, is denoted by $q_{min}$, and calculated as follows:
\begin{equation}
 q_{min} = \frac{T_{\text{total}}}{d} + \frac{\Xi_{\text{transfer}} \cdot T_{\text{total}}}{3 \cdot (\Xi_{DAO} + R_{\text{DAO}})}
 \label{minQuorum}
\end{equation}
Where $d$ is the \verb|minQuorumDivisor|. This parameter's default value is $5$, but it will double in the case the quorum has not been met for over a year.
$\Xi_{DAO}$ is the amount of ether owned by a DAO and $R_{\text{DAO}}$ is the amount of reward tokens owned by this DAO, as explained in section \ref{RewardToken} (also see \verb|rewardToken| in \ref{app:DAO}). The sum $\Xi_{DAO} + R_{\text{DAO}}$ is equal to the amount of ether used to Create DAO tokens plus the rewards received or said another way, the total amount of ether a DAO has ever received.

This means, initially, a quorum of  $20\%$ of all tokens is required for any proposal to pass. In the event $\Xi_{\text{transfer}}$ equals the amount of ether a DAO has ever received, then a quorum of $53.33\%$ is required.

In order to prevent ``proposal spam'', a minimal deposit can be required to be paid when creating a proposal, which gets refunded if quorum is achieved. If quorum is not achieved, the DAO keeps the proposal deposit. The value of the proposal deposit can be changed from the default value by the DAO through another proposal.

\section{Notation}
Throughout this paper, $\Xi$ always represents an amount of ether in its base unit wei. This is defined as $1 \: \text{Wei} = 10^{-18} \: \text{Ether}$ (\cite{Wood2014ethereum}).
Similarly, DAO tokens are denoted with $T$ and always represent the amount of DAO tokens in its base unit. We define this as $10^{-16}$ DAO token.
\section{Majority robs minority attack}
Minority owner rights can be a problem in any corporate form. Minority rights may be protected or addressed by provisions in corporate governance documents or by statute or judge-made law. But many of these solutions fail because minority owners may lack voting rights or the ability to ``vote with their feet" and easily retrieve their capital. We present a solution to this problem in the DAO's code.

@@

To prevent this, the minority must always have the ability to retrieve their portion of the funds. Our solution is to allow a DAO to split into two. If an individual, or a group of token holders, disagree with a proposal and want to retrieve their portion of the ether before the proposal gets executed, they can submit and approve a special type of proposal to form a new DAO. The token holders that voted for this proposal can then split the DAO moving their portion of the ether to this new DAO, leaving the rest alone only able to spend their own money.

This idea originates from a blog post by Vitalik Buterin (\cite{Vitalik2015subjectivity}).

A problem this simple fix doesn't address is voter apathy: some token holders might not be actively involved in their DAO and might not follow proposals closely.
An attacker could use this to their advantage. Even though the minority has the chance to retrieve their funds and split the DAO, some could be unaware of the situation and fail to act.
For a DAO to be considered safe, it is required that inactive token holders must also be protected from losing their ether. Our proposed solution is implemented by limiting each individual DAO to a single Curator.
This Curator controls the list of addresses that can receive ether from the DAO, across all proposals.
This gives the Curator of a DAO considerable power. To prevent the abuse of this power, it is possible for a DAO to vote for a new Curator, which may result in a split of the DAO as described above.

Any token holder can make a proposal to vote for a new Curator. In effect, even a single token holder is able to both retrieve their remaining portion of ether and maintain their right to any future rewards associated to their previous contribution, as these will be sent to the new DAO automatically. Rewards are defined as any ether received by a DAO generated from products the DAO funded so far and are explained in further detail in section \ref{RewardToken}.

The process of choosing a new Curator is as follows:
Any token holder can submit a proposal for a new Curator. In this case, no proposal deposit is required, because an attacker could vote for an extremely high deposit, preventing any splits. The debating period for this proposal is $7$ days.
This is $7$ days less than the minimum required for regular proposals, allowing anyone to retrieve their funds before a potentially malicious proposal goes through. There is no quorum requirement, so that every token holder has the ability to split into their own DAO. 
The debating period is used to discuss (on or off-chain) the new Curator and conduct a first vote that's non-binding. After this first vote, token holders can confirm its results or not. In the case a majority opts to  keep the original Curator, the minority can either keep the original Curator in order to avoid a split, or inversely they can confirm their vote for a new Curator and move their portion of the ether to a new DAO.

\section{Token Price} \label{TokenPrice}
DAO Token Creation schedule increases over time. This reflects an assumption that early acts of DAO Creation have greater risk, as they may have less information about the potential success of the DAO and do not know whether what contribution will lead to full funding of the DAO. 
For the DAO described here, we have chosen the following Creation schedule:
\begin{equation}
 P(t) = \begin{cases}
1 & \text{if} \quad t < t_c - 2 \cdot \text{w} \\
1 + 0.05 \cdot m(t) & \text{if} \quad t_c - 2 \cdot \text{w} \leqslant t < t_c - 4 \cdot \text{d} \\
1.5 & \text{otherwise}
\end{cases}
\end{equation}
with the multiplier $m$ defined as:
\begin{equation}
 m(t) = (t - (t_c - 2 \cdot \text{w})) / \text{d}
\end{equation}
Here $t$ is the unix time in seconds, $t_c$ is the closing time of the funding period (see \ref{app:TokenSale} \verb|closingTime|), $w$ is a week in seconds and $d$ a day in seconds.
Hence the number of tokens (in its base unit) each buyer receives is calculated as: $P(t) \cdot \Xi_c$. Here $\Xi_c$ stands for the amount of ether sent to fuel the DAO, denoted in wei.
This results in a constant Creation ratio in the beginning, until $2$ weeks before the end of the DAO Creation. At this time the Creation ratio increases daily by $0.05 \: \Xi_c$ per base unit of DAO token. Until $4$ days before the closing time when there will be a constant Creation ratio of $1.5 \: \Xi_c$ per base unit of DAO token.

Creation ratio increases during the Creation Phase could lead to a situation where a single contributor, having bought tokens during at the initial Creation ratio, could split the DAO immediately after the sale ends, thereby receiving more ether than they put in due to other contributors having a higher Creation ratio (\cite{GriffDiscussion}).
In order to avoid that possibility, all ether that is received above the initial Creation ratio of the token, will be sent to an extra account. Denoted as \verb|extraBalance| in \ref{app:TokenSale}. This ether can be sent back to the DAO through a proposal after the DAO has spent at least this amount of ether.
This rule is implemented in the internal function \verb|isRecipientAllowed| in section \ref{DAO}.

@@

The address \verb|curator| is set at the creation of the DAO and defines the Curator.

@@

The integer \verb|proposalDeposit| specifies the minimum deposit to be paid in wei for any proposal that does not include a change of the Curator.

@@

A single proposal has the parameters:
\begin{description}
 \item[recipient] The address where the \verb|amount| of wei will go to if the proposal is accepted.
 \item[amount] The amount of wei to transfer to \verb|recipient| if the proposal is accepted.
 \item[description] A plain text description of the proposal.
 \item[votingDeadline] A unix timestamp, denoting the end of the voting period.
 \item[open] A boolean which is false if the votes have already been counted, true otherwise.
 \item[proposalPassed] A boolean which is true if a quorum has been achieved with the majority approving the proposal.
 \item[proposalHash] A hash to check validity of a proposal. Equal to \verb|sha3(_recipient, _amount, _transactionData)|.
 \item[proposalDeposit] The deposit (in wei) the creator of a proposal has send to submit a proposal. It is taken from the \verb|msg.value| of a \verb|newProposal| call; its purpose is to prevent spam. Its minimal value is set when the contract is deployed as constructor parameter. But the creator of the proposal can send any amount above this for the deposit. The proposals will be sorted by the proposal deposit in the GUI, so in the case that a proposal is considered important, the creator of the proposal can deposit extra ether to “advertise” their proposal. The creator of the proposal will be refunded the entire deposit if quorum is reached, if it is not reached the deposit remains with the DAO. 
 \item[newCurator] A boolean which is true if this proposal assigns a new Curator.
 \item[splitData] The data used to split the DAO. This data is gathered from proposals when they require a new Curator.
 \item[yea] Number of tokens in favor of the proposal.
 \item[nay] Number of tokens opposed to the proposal.
 \item[votedYes] Simple mapping to check if a token holder has voted for it.
 \item[votedNo] Simple mapping to check if a token holder has voted against it.
 \item[creator] The address of the token holder that created the proposal.
\end{description}

@@

\subsubsection*{newProposal}
This function is used to create a new proposal. The arguments of the function are:
\begin{description}
 \item[recipient] The address of the recipient of the ether in the proposal (has to be the DAO address itself, the current Curator or an address on the whitelist \verb|allowedRecipients|).
 \item[amount] The amount of wei to be sent in the proposed transaction.
 \item[description] A string describing the proposal.
 \item[transactionData] The data of the proposed transaction.
 \item[debatingPeriod] The amount of time to debate the proposal, at least 2 weeks for a normal proposal and at least 1 week for a new Curator proposal.
 \item[newCurator] A boolean defining whether this proposal is for a new Curator or not.
\end{description}
After checking the sanity of the proposal (see code), this function creates a proposal which is open for voting for a certain amount of time. The function will return a proposal ID which is used to vote.

@@

\subsubsection*{executeProposal}
This function can be called by anyone. It counts the votes, in order to check whether the quorum is met, and executes the proposal if it passed, unless it is a proposal for a new Curator, than it does nothing.
The arguments of the function are:
\begin{description}
 \item[proposalID] The proposal ID.
 \item[transactionData] The data of the proposed transaction
\end{description}
The function checks whether the voting deadline has passed and that the \verb|transactionData| matches the proposal ID. Then it checks whether the quorum has been met (see Eq. \ref{minQuorum}) and if the proposal had a majority of support. If this is the case, it executes the proposal and refunds the 
proposal deposit. If the quorum has been achieved, but the proposal was declined by the majority of the voters, the proposal deposit is refunded and the proposal closes.

\subsubsection*{emergenyCloseProposal}
This is used in the case `executeProposal` fails because it throws (e.g. when returuning the deposit, because the fallback function of this address throws, or when executing the proposed transaction). It can be called $5$ days after the voting deadline of the proposal passed. It will close the proposal.
\subsubsection*{splitDAO}
After a new Curator has been proposed, and the debating period in which the token holders could vote for or against the proposal has passed, this function is called by each of the DAO token holders that want to leave the current DAO and move to a new DAO with the proposed new Curator. This function creates a new DAO and moves a portion of the ether, as well as a portion of the reward tokens to the new DAO.
The arguments are:
\begin{description}
 \item[proposalID] The proposal ID.
 \item[newCurator] The address of the new Curator of the new DAO.
\end{description}
After a sanity check (see code), this function will create the new DAO if it hasn’t already been created using the contract \verb|daoCreator|, updates the split data stored in the proposal and stores the address of the new DAO in the split data.
This function moves the portion of ether that belongs to the caller of this function in the original DAO to the new DAO. This ether amount is denoted by $\Xi_{\text{sender}}$, stated in wei and is calculated as follows:
\begin{equation}
 \Xi_{\text{sender}} = \Xi_{\text{DAO}} \cdot T_{\text{sender}} / T_{\text{total}}
\end{equation}
Here $T_{\text{sender}}$ is the amount of tokens of the caller of the function and $\Xi_{\text{DAO}}$ is the balance of the DAO at the time of the split.
This will be used to effectively buy tokens of the newly created DAO and fund the new DAO just as the original DAO was funded.
In addition to the ether which is moved to the new DAO, the reward tokens $R_{\text{sender}}$ are also transferred. 
They are calculated as follows:
\begin{equation}
 R_{\text{sender}} = R_{\text{DAO}} \cdot T_{\text{sender}} / T_{\text{total}}
\end{equation}
Where $R_{\text{DAO}}$ is the amount of reward tokens owned by the original DAO at the time of the split. These tokens allow the new DAO to retrieve their portion of the reward using the \verb|retrieveDAOReward| function of the original DAO.
At the end of this process all original DAO tokens of the sender account are destroyed. It is important to notice that in all integer division descirbed above, there may be remainders which stay with the DAO.

\subsubsection*{newContract}
This function can only be called by the DAO itself (through a proposal and a vote) and is used to move all remaining ether, as well as all rewardTokens to a new address. This is used to update the contract. The new address needs to be approved by the Curator.

@@

\subsubsection*{changeAllowedRecipients}
This function can add/remove an address to/from the whitelist, \verb|allowedRecipients|. It can only be executed by the Curator.

@@

Let’s assume that during the split, a fraction of DAO tokens, $X$, changes the Curator and leaves the DAO. The new DAO created receives $X \cdot \Xi_{\text{DAO pre}}$, a portion of the remaining ether from the original DAO.
\begin{equation}
 \Xi_{\text{DAO post}} = (1 - X) \cdot \Xi_{\text{DAO pre}}
\end{equation}
Here $\Xi_{\text{DAO pre}}$ is the ether balance of the original DAO before the split and $\Xi_{\text{DAO post}}$ is the ether balance of the original DAO after the split.

@@

The original DAO tokens of the accounts that confirmed the new Curator are destroyed.
Hence:
\begin{equation}
 T_{\text{total post}}= (1 - X) \cdot T_{\text{total pre}}
\end{equation}

@@

    // Address of the Curator
    address public curator;
    // The whitelist: List of addresses the DAO is allowed to send money to
    mapping (address => bool) public allowedRecipients;
